Deployable Probabilistic Programming

### 2019-04-17 paper/paper.tex
in progress

Incorporating a probabilistic program, or rather a probabilistic
procedure, within a larger code body appears to be rather
straightforward: one implements the model in the probabilistic
programming language, fetches and preprocesses the data in the
host programming language, passes the data and the model to an
inference algorithm, and post-processes the results in the
host programming language again to make algorithmic
decisions based on inference outcomes. However, complex
server-side software systems make integration of probabilistic
inference challenging.

@@

\paragraph{Data interface} In a server-side application data
for inference comes from a variety of sources: network,
databases, distributed file systems, and in many different
formats. Efficient inference depends on fast data access and
updating. Libraries for data access and manipulation are
available in the host environment. While the host environment
can be used as a proxy retrieving and transforming the data,
such as in the case of Stan~\cite{Stan17} integrations,
sometimes direct access from the probabilistic code is the
preferred option, for example when the data is streamed or
retrieved conditionally.

\paragraph{Integration and deployment} Deployment of server-side
software systems is a delicate process involving automatic
builds and maintenance of dependencies. Adding  a component,
which possibly introduces additional software dependencies or
even a separate runtime, complicates deployment. Minimizing the
burden of probabilistic programming on the integration
and deployment process should be a major consideration in
design or selection of probabilistic programming tools.
Probabilistic programming systems that are implemented or 
provide an interface in a popular programming language, e.g.
Python (Edward~\cite{THS+17}, Pyro~\cite{BCJ+19}) are easier
to integrate and deploy, however the smaller the footprint
of a probabilistic system, the easier is the adoption.

@@

Based on the experience of developing and deploying solutions
using different probabilistic environments, we came up with
guidelines to implementation of a probabilistic programming
facility for server-side applications. We believe that these
guidelines, when followed, help easier integration of
probabilistic programming inference into large-scale server-side
software systems.

@@

\subsubsection{Optimization}

\subsubsection{Full Posterior}

@@

A probabilistic programming facility similar to Infergo can be
added to most modern general programming languages, in
particular, those used for implementation of large-scale software
systems, making probabilistic programming inference more
accessible in server-side applications. We described Infergo
design choices and implementation techniques in intent to
encourage broader adoption of probabilistic programming in
production software systems and ease implementation of similar
facilities in other languages and software environments.

>>>>

Incorporating a probabilistic program, or rather a probabilistic
procedure, within a larger code body appears to be rather
straightforward: one implements the model in the probabilistic
programming language, fetches and preprocesses the data in the
host programming language, passes the data and the model to an
inference algorithm, and post-processes the results in the
host programming language again to make algorithmic
decisions based on inference outcomes. However, complex
software systems make integration of probabilistic
inference challenging.

@@

\paragraph{Data interface} The data for inference may come from
a variety of sources: network, databases, distributed file
systems, and in many different formats. Efficient inference
depends on fast data access and updating. Libraries for data
access and manipulation are available in the host environment.
While the host environment can be used as a proxy retrieving and
transforming the data, such as in the case of Stan~\cite{Stan17}
integrations, sometimes direct access from the probabilistic
code is the preferred option, for example when the data is
streamed or retrieved conditionally.

\paragraph{Integration and deployment} Deployment of software
systems is a delicate process involving automatic builds and
maintenance of dependencies. Adding  a component, which possibly
introduces additional software dependencies or even a separate
runtime, complicates deployment. Minimizing the burden of
probabilistic programming on the integration and deployment
process should be a major consideration in design or selection
of probabilistic programming tools.  Probabilistic programming
systems that are implemented or provide an interface in a
popular programming language, e.g.  Python
(Edward~\cite{THS+17}, Pyro~\cite{BCJ+19}) are easier to
integrate and deploy, however the smaller the footprint of a
probabilistic system, the easier is the adoption.

@@

Based on the experience of developing and deploying solutions
using different probabilistic environments, we came up with
guidelines to implementation of a probabilistic programming
facility for server-side applications. We believe that these
guidelines, when followed, help easier integration of
probabilistic programming inference into large-scale 
software systems.

@@

@@

A probabilistic programming facility similar to Infergo can be
added to most modern general programming languages, in
particular, those used for implementation of large-scale software
systems, making probabilistic programming inference more
accessible in large scale server-side applications. We described
Infergo design choices and implementation techniques in intent
to encourage broader adoption of probabilistic programming in
production software systems and ease implementation of similar
facilities in other languages and software environments.
